title: "lab14"
author: R-Kelly
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
```
# Team section
### Our Question
Our team decided to investigate the relationship between smoking, age and general health.  It is interesting and important to explore this relationship as it can give a better understanding to the effects of smoking. This answer will be interesting and important as it may lead to a better general health for the community. 
PLOT:
```{r}
ggplot(data  = cdc) + 
    geom_boxplot(mapping = aes(x  = genhlth, y = age)) +
    facet_wrap(~ smoke100, nrow = 2)
```



### Our Recommendation


We can see form the graph that people who smoke less have a higher chance of having a good health even in their older ages. So we think smoke does affect health in some extent. We recommend no smoking is the best.
# Individual Plots

### Lauren
I chose to investigate the relationship between general health and age, and the differences of this relationship for men and women. I found that females we're on average slightly healthier than men independed of age, and people from age 30 to 50 we're the heathiest group of people. 

```{r}
ggplot(data = cdc) +  geom_histogram(bins=20, mapping = aes(x = age, fill = genhlth)) +facet_wrap(~gender) +
  xlab("Age") +
  ylab("People") +
  labs(title= "General Health by Age for both Men and Women")


```

### Ping:
Question: Do people who do exercises, having health plan, no smoking, maintain a better weight? Maintaing a better weight means having a less different between their actual weight and desired weight. 

Ho: If people who do exercises, have health plan, dont smoke, then there still a difference between their actual weight and desired weight.
H1: If people who do exercises, have health plan, dont smoke, there is no difference between their actual weight and desired weight.

```{r}
perm_cor <- function(perms = 1000, weight, wtdesired)
{
 vector <- c(rep(0,perms))

 # Loop throught number of permutations
 for (i in c(1:perms))
 {
   # Step 2:
   # Randomly mix up the values in the vector "y"
   wtdesired <- sample(wtdesired, length(wtdesired))

   # Step 3:
   # Compute the correlation between x and the randomly mixed
   # up y-vector. Store this value in the vector from step 1.
   vector[i] <- cor(wtdesired, weight)
 }

 # Step 4:
 # Return new updated vector, created in step 1
 return(vector)
}
```

```{r}
cdc <- cdc %>% mutate(diff = wtdesire - weight)
cdc_training <- cdc %>% filter(exerany == "1", smoke100 == "0", hlthplan == "1")

ggplot(data = cdc_training, mapping = aes( x =diff)) + geom_histogram(color = "black", fill = "pink")

correlation <- perm_cor(perms = 1000, cdc_training$weight, cdc_training$wtdesire)
correlation <- data.frame(correlation)
cor <- cor(cdc_training$weight, cdc_training$wtdesire)

#true_value = mean(cdc$diff)
ggplot(correlation) + geom_density(aes(x = correlation)) + geom_vline(aes(xintercept=cor), linetype="dashed", size=1)


```


From the first graph, we can see that most people who do exercises, have health plan, and dont smoke have 0 difference between their actual weight and their desire weight. From the second graph, we can see that the actual correlation is way away from the graph, so it is very significant. So we can confidently reject the null. Thus if people who do exercises, have health plan, dont smoke, there is no difference between their actual weight and desired weight.


### Peter
I chose to examine the correlation between smoking, height, and weight. Since healthy weight is dependent on a number of factors, I figured the best measure of general fitness is the difference between actual and desired weight. The graph is separated by gender and smoking status, to show the difference in health between populations. 
```{r, warning = FALSE, message = FALSE}
pcdc <- cdc
pcdc <- pcdc%>% mutate(wdelta = weight-wtdesire)
valueAssign <- function(string){
  val <- 0
  if(string == "excellent"){
    val <- 5
  }
  else if(string == "very good"){
    val <- 4
  
  }
  else if(string == "good"){
    val <- 3
  }
  else if(string == "fair"){
    val <- 2
  }
  else if(string == "poor"){
    val <- 1
  }
  else{
    val <- -1
  }
  return(val)
}
pcdc <- pcdc %>% mutate(inthlth = map(genhlth, valueAssign))

ggplot(pcdc, aes(x = height, y = wdelta, color = genhlth)) +
     geom_point() +
    
    facet_grid(gender ~ smoke100)
    
```


A trend shown by the graph is the higher likelyhood of un-fit smokers being of fair or poor health. Non-smoking populations tend to have more individuals in the overweight range who's health is fair or good. Another notable trait is the higher than average health of the tallest members of each subgroup. Lastly, there is a clear trend between smoking and being of poor health. What a surprise. 



## Lauren: 

```{r}
cdclolo<- cdc %>% mutate(weightdiff = wtdesire - weight)
cdcl<- cdclolo %>% select(weightdiff, exerany, gender)
perm_mean <- function(perms)
{
  vec <- numeric(1000)
  for (i in c(1:perms))
  {
    samp <- sample(cdcl$weightdiff,2, replace = FALSE)
    value1<- samp[[1]]
    value2<- samp[[2]]
  
    
    diff = value1 - value2
    vec[[i]] <- diff
    
  }
  return(vec)
}
```

```{r}
test<-perm_mean(1000)
test<- data.frame(test)
med<-test %>% summarise(med = median(test))
med<- as.numeric(med)
ggplot() + geom_density(test,mapping = aes(x=test,stat(count))) + geom_vline(xintercept = med,color = 'red') + theme_light()
```

```{r}
fcdc<- cdcl %>% filter(gender == 'f') %>% summarise(mean(weightdiff))
exerf<- cdcl %>% filter(gender == 'f', exerany == 1)%>% summarise(mean(weightdiff))
nexerf<- cdcl %>% filter(gender == 'f', exerany == 0)%>% summarise(mean(weightdiff))

mcdc<- cdcl %>% filter(gender == 'm') %>% summarise(mean(weightdiff))
exerm<- cdcl %>% filter(gender == 'm', exerany == 1)%>% summarise(mean(weightdiff))
nexerm<- cdcl %>% filter(gender == 'm', exerany == 0)%>% summarise(mean(weightdiff))
```


I chose to investigate the relatioship bewteen the difference between desired weight loss and current weight, and if the person exersized. After calculating the difference in disired weight and current weight, it became clear that many people wanted to gain weight, just as many people wanted to loose weight. To see if there was any correlation between people going to the gym, and wanting to gain or loose weight I created a hypothesis test where I ignored the exerany data and choose two data points at random to calculate the difference in 'diffweight'. After running the hypothesis test, it is clear that the mean diffweight is 0, and in fact it does not depend on if the person is exesizing at all. 

To investigate any possible confounding variables, I ran a series of calculations on groups of males and females to see their average weightdiff. On average, women wanted to loose 8 more pounds than men. Out of those who had worked out in the last month, women on averagae wanted to loose 7 more pounds than men, and out of those who had not worked out in the last month, women wanted to loose 9 pounds more on average than men. 

To conclude, more men are looking to gain weight than women, which could be a confounding variable to my original findings. 

## Sasha:

```{r}
library(modelr)
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
```


```{r}

options(na.action = na.warn)

cdcN<- na.omit(cdc)

cdc2 <- lm(wtdesire ~ weight, data = cdcN)
coef(cdc2)

cdcN <- cdcN %>% 
  add_residuals(cdc2)

ggplot(data = cdcN, aes(x = wtdesire, fill = gender))+
  geom_density() +
  facet_wrap("genhlth") +
  xlim( 0 , 300)+
  geom_vline(aes(xintercept = mean(wtdesire)), color = "black", linetype = "solid", size = 1)+
  xlab("Weight Desire")+
  ylab("Density")

ggplot(data = cdcN, aes(x = resid, fill = gender))+
  geom_density()+
  facet_wrap("genhlth")+
  xlim(-100, 100)+
  geom_vline(aes(xintercept = mean(resid)), color = "black", linetype = "solid", size = 1)+
  xlab("Residuals")+
  ylab("Density")

```
I chose to examine the relationship between weight, weight desire and how those are 
affected by gender and general health. Through the use of a linear model I was able 
to graph the residuals of weight desire and actual weight. This reveals that people 
who are in better health tend to be closer to their ideal wight and as health levels 
decrease, people tend to be further from their ideal weight. Another interesting 
revelation presented in these graohs is that women tend to to want to weigh less than 
they do while men tend to want to weigh more.



## Gregor:

```{r}
GregorData <- cdc
#renaming col names to be more clear what they are
colnames(GregorData)[1] <- "GeneralHealth"
colnames(GregorData)[2] <- "HasExercised"
colnames(GregorData)[3] <- "HasHealthplan"
colnames(GregorData)[4] <- "HasSmoked100Cigs"

quantify <- function(healthCondition){
    if (healthCondition == "poor"){
      return(1)
    }

    else if (healthCondition == "fair"){
      return(2)
    }
  
    else if (healthCondition == "good"){
      return(3)
    }
  else if(healthCondition == "very good"){
      return(4)
  }
  
  else if (healthCondition == "excellent"){
      return(5)
  }
}




GregorData <- GregorData %>% mutate(GenHealthQuantify = as.numeric(map(GeneralHealth, quantify)))
GregorData <- GregorData %>% mutate(WeightDesiredDiff = wtdesire - weight )
GregorData <- GregorData %>% mutate(isSatisfied = ifelse(wtdesire == weight, TRUE, FALSE))
GregorData <- GregorData %>% mutate(wantsBigger = ifelse(wtdesire - weight > 5, TRUE, FALSE))




```

```{r, include=FALSE, echo = TRUE}
ggplot(GregorData, aes(x = weight, y= WeightDesiredDiff)) + geom_point() + facet_wrap(~gender)

```

```{r, include = FALSE, echo = TRUE}
#include = FALSE won't show it, as this isn't final graph but just observing relationships, but code will show
ggplot(GregorData, aes(x = GeneralHealth)) + geom_histogram(stat = "count", fill = "steel blue") + facet_wrap(~isSatisfied)
#not enough of a difference between those who don't care about weight and health, although it is interesting that those that are satisfied with weight have about even good health, and greater proportion of excellent. Could be they don't care, or they healthy with weight. nonetheless not very strong relationship


```

```{r, include = FALSE, echo = FALSE}
#those who desire to be bigger don't show any different health states, thinkin that guys that want to bulk up and start going to the gym would maybe be healthier, but guess not
ggplot(GregorData, aes(x = GeneralHealth)) + geom_histogram(stat = "count", fill = "dark red") + facet_grid(gender~wantsBigger) 
```


```{r}

GregorModel <- lm(as.numeric(GenHealthQuantify) ~ WeightDesiredDiff, data= GregorData)
GregorData <- add_residuals(GregorData, GregorModel)
ggplot(GregorData, aes(x = resid)) + geom_density(bins = 50, fill = "steel blue") + facet_wrap(~HasExercised) + ylim(0, 1) + ggtitle("The difference in Residuals of those that exercised with those that don't")
```


The residual on this graph is the difference between the model of a quantified version of general health and each person's desired weight difference. General Health ranges from 1-5, with 1 representing poor health, and with 5 representing excellent health. The desired weight difference is a variable that represents how much they wish they're weight was different, with negative values representing that person wishes he/she is that many pounds lighter, and with positive values representing that person wishes he/she is that many pounds heavier. Therefore, the residual represents how far off each person is from this model. 

The graph shows residuals and count, with a facet wrap of those who have exercised in the past month on the right (1), and those that haven't exercised in the past month. In the group that has exercised more, there's a much higher proportion of people following the model more closely, with the greatest density of data being around the 0.0 mark. Hence, those that exercise more closely follow the model of those that have a greater desire weight difference have a greater general health, which could make sense, considering those that care about their weight will genereally be healthier. 






The graphs show that whether one exercises has a bigger impact on one's general health. In the graphs that show the people who have exercised in the past 
month (meaning of the 'exerany' variable), the ratio between excellent and very good status to fair and poor is much greater than the same ratio in the graphs of people
who don't smoke. The fact that this ratio is still much greater in the people who smoke and exercise in comparison to this same ratio in the people who smoke but don't 
exercise demonstrates that exercise has a bigger impact on general health. In the two graphs that show the people that don't smoke, the ratios between good 
health and poor health is about the same, whether the person smokes or not. 

```{r}
ggplot(data  = cdc, mapping = aes(x  = genhlth)) + geom_bar() + facet_wrap(exerany~smoke100, labeller = label_both)
``` 
#Who Did What!

Lauren: In this team Lab, I worked on formatting the document and fixing code so all of the plots would print when the file was knit. I also worked on developing and writing the team question for our plot. 
Ping: I did the team plot and recommendations.





