---
title: "Lab10"
author: "Gregor Tzinov"
date: "March 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```




```{r}
#dont include your own code, lets keep them up here so that we don't have the program load libraries twice
library(tidyverse)
library(dplyr)
```


```{r}
answers_trunc <- read.csv("Answers_trunc.csv")
questions_trunc <- read.csv("Questions_trunc.csv")
```

#Team

From all of our individual data manipulation and investigations, we found the following about question and answers:
    1. There is no correlation between the length of the question asked, and how long it takes to answer it
    2. Very long/very short answers do not get rated very highly
    3. Personal questions have a lower average rating than more general questions
    4. The best questions and answers are asked at 6pm and 10pm.
    
According to our findings, the ideal questions and answers are ones which aren't too short, nor too long. However, shorter questions dont necessarily get an answer faster. If one is looking for an answer quickly, they don't need to wory about the length of the question. It is also beneficial to ask questions at 6:00pm or 10:00pm. 

# Individual Sections


### Gregor

```{r}
gregor_answers_trunc <- mutate(answers_trunc, ContainsQuestion = str_detect(answers_trunc$Body, "\\?"), ContainsTry = str_detect(answers_trunc$Body, "try"), Length400_600 = (str_length(answers_trunc$Body) > 400 & str_length(answers_trunc$Body) < 600))
ggplot(data = gregor_answers_trunc, aes(x = Score)) + geom_histogram(bins = 100) + facet_grid(ContainsQuestion~Length400_600, labeller = label_both) + xlim(0, 100) + ylim(0, 1000) + xlim(-100, 100)

```

I wanted to determine whether or not there was a relationship between answer length and whether the answer had a question mark with what the rating of the answer is. The data shown above can lead to various conclusions. The filter combination that lead to the greatest average score was when the answer did not contain the question mark, nor was the answer in between 400 - 600 characters. What this could represent is that people do not rate answers that are very long/very short highly. However this filter combination also has the highest proportion to negative scores, which could be caused by those answers that are very, very short, such as only a few characters, that people will clearly rate poorly generally. The filter that has the lowest average score is the group of answers that is between 400 and 600 characters, and there's a question mark included. People tend to not like being asked questions back in an answer that has moderate length. Perhaps when people are reading this response, they want to spend time to find exactly what they're looking for, rather than being asked a counter question.

### Peter

```{r}
dat <- questions_trunc
dat$myInT <- grepl("my", dat$Title, fixed = TRUE)
ggplot(data = dat, mapping = aes(x = Score, y = as.factor(myInT), color = Score)) + geom_jitter() + ylab("If Question contains the word 'My'")
```
This graph shows the score relationship between questions whose titles do or do not contain the character string "my". The presence of this string indicates a personal question, which is likley to be a program specific problem, which is unlikley to be helpful to other people with similar issues.  
This assertion is supported by both the graph, and the mean values for score in each category: Questions containing 'my' have a mean score of 13.47, while questions without 'my' have a mean score of 25.69. 



### Ping
```{r}
ping <- mutate(answers_trunc,Pythonanswer = str_detect(answers_trunc$Body, "Python")) 
ggplot(data = ping, aes(x = Score)) + geom_histogram(bin = 10, col = "blue", main = "Histogram for scores distribution for Python related questions") + xlim(0, 100)+  ylim(0,1000) +
  facet_wrap(~Pythonanswer)
```
I want to see how frequent people ask about Python related questions on this website and how people are satisfied with the answers related with Python questions. From the grpah, we can see that almost equally, people ask about Python related questions versus non python related questions. And we can also see that the scores for both Python related questions or non Python related questions, distribute mainly from 5 -50.  



### Sasha

```{r}
sanswers_trunc <- answers_trunc 
squestions_trunc <- questions_trunc 
sanswers_trunc$X <- NULL
squestions_trunc$X <- NULL

library(stringr)
library(tidyverse)
library(dplyr)

squestions_trunc <- squestions_trunc %>% 
  separate(CreationDate, into = c("Date", "Time"), sep = "T") 

squestions_trunc <- squestions_trunc %>% 
  separate(Time, into = c("Hours", "Minutes", "Seconds"), sep = ":")

squestions_trunc$Seconds <- NULL 
squestions_trunc$Minutes <- NULL 

sanswers_trunc <- sanswers_trunc %>% 
  separate(CreationDate, into = c("Date", "Time"), sep = "T") 
sanswers_trunc <- sanswers_trunc %>% 
  separate(Time, into = c("Hours", "Minutes", "Seconds"), sep = ":") 

sanswers_trunc$Seconds <- NULL 
sanswers_trunc$Minutes <- NULL 

ggplot(data = sanswers_trunc, mapping = aes(x = as.factor(Hours), y = Score)) +
  geom_bar(stat="identity", color = "lightpink", fill = "lightpink")+
  xlab("Hour of the Day")+
  ggtitle("Score of Answers by Hour")

ggplot(data = squestions_trunc, mapping = aes(x = as.factor(Hours), y = Score)) +
  geom_bar(stat="identity", color = "lightblue", fill = "lightblue")+
  xlab("Hour of the Day")+
  ggtitle("Score of Questions by Hour")


```

I chose to explore the relationship between Score and time of day for both questions 
and answers. I seperated the CreationDate column into date and time and then further 
seperated time by hours, minutes and seconds. I then nullified seconds to narrow the 
scope to just hours. From there, I plotted in ggpolt using a barplot to show the score 
for questions and answers per hour. From the graph, it's seen that questions asked at 
6 pm and 10 pm have the greatest score which lends itself to the idea that those questions
 are deemed more pertinent and will likely recieve more attention and therefore better 
 answers. On the answer side there is similar result with the highest scores occuring at 6 pm
 and again at 10 pm.

### Lauren

```{r}
answers <- answers_trunc
answers$Id <- answers$ParentId
answers$AnswerDate <- answers$CreationDate
answers <- answers %>% select(Id,AnswerDate)

questions <- questions_trunc
questions$QuestionDate <- questions$CreationDate
questions <- questions %>% select(Id,QuestionDate,Body)
qa<- full_join(questions,answers,by = "Id")
qa <- qa %>% separate(QuestionDate,into= c("QDate","QTime"), sep="T")
qa <- qa %>% separate(AnswerDate, into = c("ADate", "ATime"),sep = "T")

qa <- qa %>% separate(QDate,into= c("QYear","QMonth","QDay"), sep="-")
qa <- qa %>% separate(QTime, into = c("QHour", "QMin","QSecond"),sep = ":")
qa <- qa %>% separate(ADate,into= c("AYear","AMonth","ADay"), sep="-")
qa <- qa %>% separate(ATime, into = c("AHour", "AMin","ASecond"),sep = ":")
qa$QSecond <- substr(qa$QSecond,1,nchar(qa$QSecond)-1)
qa$ASecond <- substr(qa$ASecond,1,nchar(qa$ASecond)-1)

qa<- qa %>% na.omit(qa)

qa$QYear<- as.numeric(qa$QYear)
qa$QMonth<- as.numeric(qa$QMonth)
qa$QDay<- as.numeric(qa$QDay)
qa$QHour<- as.numeric(qa$QHour)
qa$QMinute<- as.numeric(qa$QMin)
qa$QSecond<- as.numeric(qa$QSecond)

qa$AYear<- as.numeric(qa$AYear)
qa$AMonth<- as.numeric(qa$AMonth)
qa$ADay<- as.numeric(qa$ADay)
qa$AHour<- as.numeric(qa$AHour)
qa$AMinute<- as.numeric(qa$AMin)
qa$ASecond<- as.numeric(qa$ASecond)

qa$QDT<- ISOdatetime(qa$QYear,qa$QMonth,qa$QDay,qa$QHour,qa$QMin,qa$QSecond)
qa$ADT<- ISOdatetime(qa$AYear,qa$AMonth,qa$ADay,qa$AHour,qa$AMin,qa$ASecond)
qa<- qa%>%select(QDT,ADT,Id,Body)
qa<- qa %>% mutate(timediff = difftime(ADT,QDT,units = "hours"))
qa$Body <- as.character(qa$Body)
qa<- qa %>% mutate(qlength = nchar(Body))

ggplot(qa)+geom_jitter(aes(x=timediff,y=qlength,color=),alpha = 1/5,color="hot pink")+coord_cartesian(xlim = c(0, 10000), ylim = c(0, 10000))

qa$timediff<- as.numeric(qa$timediff)
qa$qlength<- as.numeric(qa$qlength)
corr<-cor(qa$timediff,qa$qlength)
print(corr)
```

I chose to investigate the relationship between the time it took to answer a question, and the length of the question. I did so by creating a new column which calculated the difference in the time a question was asked and asnwered. I then created another column which calculated the length of the question. According to the plot, and the correlation calculation, there is almost no correlation between the time it takes to answer a question and the length of the question. The reason there was no correlation could be answered by taking a further look into the relationship between the question topic and the time it took to answer a question. 
