

---
title: "Lab 12"
author: "Team RKelly: Gregor, Ping, Sasha, Lauren, Peter"
date: "April 11, 2019"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```



```{r}
library(tidyverse)
library(dplyr)
```


```{r}

gapminder<- gapminder::gapminder
samp <- sample(gapminder$lifeExp,300, replace = FALSE)
#view(samp)
```


```{r}

```

```{r}

v1 <- gapminder$year
v2 <- gapminder$lifeExp
group1 <- samp[1:3]
group2 <- samp[3+1:300]

```



```{r}
perm_mean <- function(perms = 1000, values, n1)
{
  ## Variables ##
  # perms: The number of permutations 
  # values (num): 
  # n1 (int): Size of group 1
  ###############
  vec <- numeric(1000)
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  
  
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly separate vector "values" into disjoint 
    # groups of size "n1" and "length(values) - n1" respectively
    
    samp <- sample(gapminder$lifeExp, 300, replace = FALSE)
    group1 <- samp[1:n1]
    group2 <- samp[n1+1:length(values)]
    
    # Step 3:
    # Compute the sample means for the two groups from 
    # step 2
    
    M1 = mean(group1)
    M2 = mean(group2)
    
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1
    
    diffmean = M1 - M2
    vec[[i]] <- diffmean
    
  }
  
  # Step 5:
  # Return new updated vector, created in step 1
  return(vec)
  
}
```

```{r}
test<-perm_mean(1000, 300, 50)
test<- data.frame(test)

test1 <- perm_mean(1000, 300, 50)
test2 <- perm_mean(1000, 400, 30)
test3 <- perm_mean(1000, 500, 20)

test1 <- data.frame(test1)
#test1 %>% mutate(testNum = "test1")
test2 <- data.frame(test2)
#test2 %>% mutate(testNum = "test2")
test3 <- data.frame(test3)
#test3 %>% mutate(testNum = "test3")

ggplot(test1, aes(x = test1))+
  geom_density(color = "magenta", fill = "pink", alpha = 0.6)+
  xlab("Sample Mean")+
  ylab("Mean Density")+
  ggtitle("Sample Test 1")

ggplot(test2, aes(x = test2))+
  geom_density(color = "dark blue", fill = "blue", alpha = 0.2)+
  xlab("Sample Mean")+
  ylab("Mean Density")+
  ggtitle("Sample Test 2")

ggplot(test3, aes(x = test3))+
  geom_density(color = "dark green", fill = "green", alpha = 0.2)+
  xlab("Sample Mean")+
  ylab("Mean Density")+
  ggtitle("Sample Test 3")

```


Question: Does sample size difference have a statistical relevance, and how does that affect sample mean?

The graphs above show that as we vary sample size, the distribution of Sample Mean varies. Test 1 has sample difference of 250, Test 2 has sample difference of 370, and Test 3 has sample difference of 480. As we look at the graphs, there is a greater sample with Test 3, where the sample difference is the greatest. The closer the two samples are, the closer sample mean will be to 0.


```{r}
perm_corr <- function(perms = 1000, v1, v2)
{
  ## Variables ##
  # perms: The number of permutations 
  # v1 : first correlation vector
  # v2 : second correlation vector
  ###############
  vec <- numeric(1000)
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  
  
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    mixedV2 <- sample(v2)
    # Step 3:
    # Compute the sample means for the two groups from 
    # step 2
    
    vec [[i]] <- cor(v1,mixedV2)
    
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1
    
    
    
  }
  
  # Step 5:
  # Return new updated vector, created in step 1
  return(vec)
  
}

```


```{r}

corTest <- perm_corr(1000, v1, v2)
plot(corTest)
#print(corTest)
```

Question: Is there a correlation between life expectancy and year?

The graph above shows that there is essentially no correlation between life expectancy and year. We're testing correlation between life expectancy and year with the vectors v1 and v2, which represent year and life expectancy, respectively. At all indexes, the correlation is near 0, and if it does vary, it doesn't exceed +- .05, which shows very little correlation. 



