---
title: "Lab10"
author: "Gregor Tzinov"
date: "March 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```




```{r}
#dont include your own code, lets keep them up here so that we don't have the program load libraries twice
library(tidyverse)
library(dplyr)
```


```{r}
answers_trunc <- read.csv("Answers_trunc.csv")
questions_trunc <- read.csv("Questions_trunc.csv")
```


# Individual Sections


### Gregor

### Peter

```{r}
dat <- questions_trunc
dat$myInT <- grepl("my", dat$Title, fixed = TRUE)
ggplot(data = dat, mapping = aes(x = Score, y = as.factor(myInT), color = Score)) + geom_jitter() + ylab("If Question contains the word 'My'")
```
This graph shows the score relationship between questions whose titles do or do not contain the character string "my". The presence of this string indicates a personal question, which is likley to be a program specific problem, which is unlikley to be helpful to other people with similar issues.  
This assertion is supported by both the graph, and the mean values for score in each category: Questions containing 'my' have a mean score of 13.47, while questions without 'my' have a mean score of 25.69. 



### Ping

### Sasha

### Lauren
