---
title: "Lab 4: Smoking and Low Birth Weight?"
author: "R-Kelly: Lauren Tafoya, Peter Bullard, Gregor Tzinov, Ping Li, and Sasha Maslowski"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) #renaming `bwt/oz` because the "/" sign requires special care
```

##Overall Questions
##Conclusions

*put plot here

##Our Recommendation
##Preliminary Question
##Individual Findings
####Lauren

```{r}
ggplot(babies,mapping=aes(x=bwtoz ,y=gestation))+geom_point()+geom_smooth()
```
####Peter
```{r, echo = FALSE}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`)
B <- split(babies, babies$smoke)
#c <- ggplot(babies, aes(as.factor(smoke), bwtoz)) 


vars <- colnames(babies)
Nsmoke <- B[[1]]
Smoke <- B[[2]]

x <- vector(mode = 'numeric', length(vars))
y <- vector(mode = 'numeric', length(vars))
for (i in 1:length(vars)){
  
  nsm <- cor(Nsmoke$gestation, Nsmoke[i], use = "complete.obs")
  sm <- cor(Smoke$gestation, Smoke[i], use = "complete.obs")
  
  x[[i]] <- nsm
  y[[i]] <- sm
  
  
  
}
c <- ggplot()
c <- c + geom_point(aes(x, as.factor(vars) ), color = 'red')
c <- c + geom_point(aes(y, as.factor(vars)), color = 'blue')
c <- c + xlab("Correlation")
c <- c + ylab("Field")
c <- c + ggtitle("Correlation with Gestation Period", subtitle = "Red = Non-Smoking, Blue = Smoking")
print(c)

```
I created a plot that shows the differences in correlation between gestation period and all other fields for smokers and non-smokers. The figure facilitates analysis of factors other than smoking,   
showing their influence on the variables we are studying. The large difference in correlation between birth weight and gestation period indicates that smoking lowers birth weight at all gestation lengths,  
because a high correlation means that both values grow together. 



####Gregor
```{r}
 ggplot(data = babies) + geom_bar(mapping = aes(x = mage, fill = as.factor(Premature) ), bins = 30) + labs(x = "Mother's Age", y = "Number of Women", fill = "Premature Birth or Not") + facet_wrap(~smoke)

 I, Gregor Tzinov, created a plot that shows how mother's age and whether she smokes or not affects whether the birth would be premature. The graphs demonstrate that as the mother grows in age, smoking will increase the chance that the 
 birth will be premature. There is more blue prominent (displaying premature births) from ages 30 - 40 in the graph that shows mothers that smoke in comparison to the graph that shows mothers who do not smoke. 

```
####Ping
####Sasha
##Who did what
Lauren:
Peter:
Gregor:
Ping:
Sasha:
