---
title: "Lab 6: Airport Data Science"
author: R-Kelly
date: 2019-02-19
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")
```

###Team Section 
##Main Questions:  
#What are the most likely factors to delay flights arriving in Denver?   
#What variables, if any, have no predictable effect on arrival delays?
##Team Plot:

##Findings:
#Peter:  
I found that flight speed has almost no effect of the arrival delay of flights, as the disrtibution of late flights matches that of flight speed. 
##Recomendation:  


###Personal Section
####Peter

This plot shows the relative distribution of flights delay times, and their average air speed. This graph shows that there is almost no correlation between arrival delay and airspeed of the flight. 
```{r}
Cflights <- mutate(COflights, distAir = DISTANCE/AIR_TIME)
c <- ggplot(data = COflights, mapping = aes(x = ARR_DELAY, y = distAir)) +
   geom_point() +
   geom_smooth()
print(c)
```


####Lauren



####Gregor



####Sasha




####Ping


#What we did:  
#Peter:  
I examined the relationship between average airspeed and arrival delay, as well as helped with the formatting and formulation of our research questions. 

#Lauren

#Gregor

#Sasha

#Ping



