---
title: 'Lab 6: Airport Data Science'
author: "R-Kelly"
date: '2019-02-19'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")
```

#Team Section 

##Main Questions:  

####What are the most likely factors to delay flights arriving in Denver?  

It is interesting to note that the probability of a Late Aircraft Delay occuring given that there is a arrival delay, is higher when looking at the number of minutes produced by each delay versus the number of actual delays. With that being said, Late Aircraft Delays, Carrier Delays, and NAS Delays are still the most likely factor to delay flights arriving to Denver. 

####What variables, if any, have no predictable effect on arrival delays?   

As seen in Peter's individual plot, we can conclude that the average air speed does not have a predictable effect on arrival delays. It is also interesting to note that weather delays and security delays are the least likely to be a factor for a delayed flight arriving to Denver. 

##Team Plot:

```{r}

landed<- COflights %>% filter(ORIGIN == "DEN", CANCELLED==0)

late<-landed %>% filter(ARR_DELAY>=15) %>% mutate(DELAY= ifelse(WEATHER_DELAY != 0,1,0) + ifelse(NAS_DELAY != 0,1,0) + ifelse(SECURITY_DELAY != 0,1,0) + ifelse(CARRIER_DELAY != 0,1,0) + ifelse(LATE_AIRCRAFT_DELAY != 0,1,0)) 

delaytype <- c("Carrier Delay", "Weather Delay", "NAS Delay", "Security Delay", "Late Aircraft Delay")

total_time_delayed <- c(summarise(late, sum(CARRIER_DELAY)),summarise(late, sum(WEATHER_DELAY)),summarise(late, sum(NAS_DELAY)),summarise(late, sum(SECURITY_DELAY)),summarise(late, sum(LATE_AIRCRAFT_DELAY)) )
delay_data <- data.frame(delaytype, total_time_delayed)
ggplot(delay_data)+geom_col(mapping = aes(x=delaytype, y=total_time_delayed), fill = 'light pink') + theme_minimal()
```

##Findings:

####Peter:  

I found that flight speed has almost no effect of the arrival delay of flights, as the disrtibution of late flights matches that of flight speed. 

####Ping:

According to the plot I showed below, we can see that the month of the year and different brands of the carriers do affect the delay of the flights a little bit. We can see that May through August are the months having the highest probability of a delayed flight. WN, UA, OO carriers have the highest probability of having a delayed flight. So, all in all, there is a higher chance of taking a delayed flight by WN, UA, and OO carriers in May through August.  

##Recomendation:  

####Ping:

When it comes to travelling, choose the months that are not that busy, for example, months from Sep to April. Also, especially try to avoid choosing WN, UA, OO carriers in busy months, because they have a higher chance of being delayed.

####Lauren: 

I would recommend to the airport managers to focus on carrier delays since this is the one variable which is most in the airline's control. The managers should work on a more effective process in this area. 

###Personal Section

####Peter

This plot shows the relative distribution of flights delay times, and their average air speed. This graph shows that there is almost no correlation between arrival delay and airspeed of the flight. 
```{r, message = FALSE, warning = FALSE}
Cflights <- mutate(COflights, distAir = DISTANCE/AIR_TIME)
c <- ggplot(data = Cflights, mapping = aes(x = ARR_DELAY, y = distAir)) +
   geom_hex() +
   geom_smooth()
print(c)
```


####Lauren

```{r echo=TRUE}
nlate<- summarise(late, sum(DELAY))

nweather<-landed %>% filter(ARR_DELAY>=15 & WEATHER_DELAY>0) %>% count()
pweather<-nweather/nlate

ncarrier<-landed %>% filter(ARR_DELAY>=15 & CARRIER_DELAY>0) %>% count()
pcarrier<-ncarrier/nlate

nnas<-landed %>% filter(ARR_DELAY>=15 & NAS_DELAY>0) %>% count()
pnas<-nnas/nlate

nsecurity<-landed %>% filter(ARR_DELAY>=15 & SECURITY_DELAY>0) %>% count()
psecurity<-nsecurity/nlate

nlateaircraft<-landed %>% filter(ARR_DELAY>=15 & LATE_AIRCRAFT_DELAY>0) %>% count()
plateaircraft<-nlateaircraft/nlate
```

```{r}
delaytype <- c("Carrier Delay", "Weather Delay", "NAS Delay", "Security Delay", "Late Aircraft Delay")
amount_of_delays <- c(20997,1667,23199,38,21087 )
delay_data <- data.frame(delaytype, amount_of_delays)
ggplot(delay_data)+geom_col(mapping = aes(x=delaytype, y=amount_of_delays)) + theme_dark()
```

From this representation of the data it is clear that the largest amount of Arrival Delays occurs due to National Air System delays, with delays due to Late Aircraft and Carrier Delays
close behind. The proabilities of each delay type occuring based on the number of delays (not the delays in minutes) is as follows. 


Carrier Delay | Weather Delay | NAS Delay | Security Delay | Late Aircraft Delay
---- | ------ | ------- | ------- | ------ 
31.3% | 2.49%   | 34.6% |  .0567% | 31.5%

####Gregor

```{r, echo = FALSE}
late_taxi_out <- filter(COflights, TAXI_OUT>= 15)
late_taxi_in <- filter(COflights, TAXI_IN >= 15)
num_late_taxi_out_delayed <- count(filter(late_taxi_out, ARR_DELAY >= 15))
num_late_taxi_in_delayed <- count(filter(late_taxi_in, ARR_DELAY >= 15))
```

Probability of flight that has delayed taxi out is delayed with arrival is:
```{r, echo=FALSE}
 num_late_taxi_out_delayed[1][1] / count(COflights) * 100
```

Probability of flight that has delayed taxi in is delayed with arrival is:
```{r, echo=FALSE}
num_late_taxi_in_delayed[1][1] / count(COflights) * 100
```

Creating vectors to contain names and values
```{r}
x = c("Probability of flights that had late taxi times on arrival being delayed", "Probability of flights that had late taxi times on departure being delayed")

y = c(num_late_taxi_in_delayed[1][1] / count(COflights) * 100, num_late_taxi_out_delayed[1][1] / count(COflights)*100)
```

Creating Tibble

```{r}
GraphData <- tibble(x, y)
```


```{r}
ggplot(data = GraphData, aes(x = x, y = y)) + geom_bar(stat = "identity") + coord_flip() + ylab("Percentage") + xlab("Type of taxi delay")
```



With these results, it is clear that taxing on DENs side doesn't affect whether the flight is delayed. However, it is unknown whether the arrival time is recorded before or after taxing. If it is recorded after taxing, then this information is significant as it shows that DEN airport is efficient in taxing, since there's a low chance that flights that are delayed is due to late taxing. However, if the arrival time is recorded before taxing, then this information is insignificant in relation to identifying what DEN airport can do to adjust for preventing delayed flights. 

When flights that have a delayed taxi time leaving the origin airport to DEN, there's about a 10% chance that they will be delayed coming into DEN. The airport could perhaps use this information as if they see a flight has a taxi time of more than 15 minutes, than they can expect that 10% of the time it will be a delayed flight, which can be used to inform people that are awaiting passsengers of that flight. 



####Sasha

```{r}
 firstHalf <- filter(COflights, DAY_OF_MONTH <= 15)
firstHalfDelayed <- count(filter(firstHalf, ARR_DELAY >= 15))
firstHalfDelayed[1][1] / count(COflights) * 100
```
```{r}
secondHalf <- filter(COflights, DAY_OF_MONTH >15)
secondHalfDelayed <- count(filter(secondHalf, ARR_DELAY >= 15))
secondHalfDelayed[1][1] / count(COflights) * 100

ggplot(data = firstHalf) + 
  geom_bar(mapping = aes(x = ARR_DELAY))+ xlim(0,60) + xlab( "Arrival Delay")

ggplot(data = secondHalf) + 
  geom_bar(mapping = aes(x = ARR_DELAY))+ xlim(0,60) + xlab( "Arrival Delay")
```





####Ping
I try to explore how can the month of the year and the brands of the carriers affect the delay of the flight to Denver.

```{r echo=TRUE}

DenDelayed_Month_Carrier <- COflights %>% filter(DEST == "DEN", ARR_DELAY >= 15) %>% select(YEAR, MONTH, CARRIER, ARR_DELAY)
ggplot(data = DenDelayed_Month_Carrier) + geom_bar(mapping = aes(x = MONTH, fill = CARRIER)) + xlim(0,13)

total_allMonths <- DenDelayed_Month_Carrier %>% count()
highlyDelayedCount <- DenDelayed_Month_Carrier %>% filter((MONTH %in% c(5, 8)))  %>% count()
prob <- (highlyDelayedCount/total_allMonths)

pro_may_aug_lowdelayed <- (DenDelayed_Month_Carrier %>% filter(CARRIER %in% c("DL", "AA", "VX")) %>% count()) / total_allMonths

pro_may_aug_highdelayed <- (DenDelayed_Month_Carrier %>% filter(CARRIER %in% c("WN", "VA", "OO")) %>% count()) / total_allMonths

```
We can see from the information above, that the probability of a delayed flight in May through August is 20.58%. In May through August, if you take the flights of WN, UA, OO, there is a even higher chance of being delayed(53.34%), compaing to taking DL, AA, VX(11.16%).


##What we did:  
###Peter:  
I examined the relationship between average airspeed and arrival delay, as well as helped with the formatting of the report and formulation of our research questions. 

###Lauren
I examined the probability of different delay types given that a flight was delayed by calculating probabilities and representing the data in a bar graph. 

###Gregor
Examined the probability of flights that have delayed taxi times delaying the actual arrival time of the flight


###Sasha:
I explored the probablity of arrival delay for flights that occur in the first half of the month (until the 15th day) and then also the second half of the month. 
According to the data, it is more probable that a flight will have an arrival delay earlier in the month as opposed to later.

###Ping:  
I explored the relationship between month, carriers and the delayed flights. I also helped with the conclusion of the Team Findings. 



