---
title: "Lab7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)
library(data.table)
library(lubridate)
powerC<-read.table('household_power_consumption.txt', sep=';',header= T, na.strings = c("NA","","?"), stringsAsFactors=FALSE)
str(data5)
```


###Lauren

```{r}
powerConsum <- powerC %>% 
  mutate(Year = year(dmy(Date)), Month = month(dmy(Date)), Day = day(dmy(Date)))

Winter <- powerConsum %>% 
  filter(Month %in% c(12,1,2))
powerconsumwinter <- summarise(Winter, sum(Sub_metering_3, na.rm=T))

Summer <- powerConsum %>% 
  filter(Month %in% c(6,7,8))
powerconsumsummer <- summarise(Summer, sum(Sub_metering_3, na.rm=T))

Spring <- powerConsum %>% 
  filter(Month %in% c(3,4,5))
powerconsumspring <- summarise(Spring, sum(Sub_metering_3, na.rm=T))

Fall <- powerConsum %>% 
  filter(Month %in% c(9,10,11))
powerconsumfall <- summarise(Fall, sum(Sub_metering_3, na.rm=T))

season <- c("Winter", "Spring", "Summer", "Fall")
power_consumption_sub_metering_3 <- c(powerconsumwinter,powerconsumspring,powerconsumsummer,powerconsumfall)
season_data <- data.frame(season, power_consumption_sub_metering_3)
ggplot(season_data)+geom_col(mapping = aes(x=season, y=power_consumption_sub_metering_3), fill = 'hot pink') + theme_classic()

```

Question: Does the use of airconditiong correlate to the season?

Answer: My plot shows the relationship between overall power used in sub metering 3, and the seasons. I first broke up the data into the four seasons, and then summed the total energy used for sub metering 3 for each season. The data showed more total energy used in winter than any of the other seasons, and the least amount of energy used during the summer months. According the the climate of the region (Sceaux, France), the winter and spring seasons are generally much colder than the fall and the summer months. This would indicate a use of air conditioning during spring and winter, and a use of heating during summer and fall. The data then shows that generally, heating the house uses more energy than air conditioning. 

Of course, one must consider the many assumptions used to come to this conclusion, and there may be error in this generalization. 

###Gregor

```{r}
data <- powerC %>%
  separate(Time, into  = c("Hours", "Minutes", "seconds"), sep = ":")

filtered_time <- filter(data, Hours == 22)

ggplot(data = filtered_time, aes(x = Minutes, y = Sub_metering_1)) + geom_jitter() + ylab("Energy use in kitchen") + xlab("Minutes past 11PM")

```

Answer: When looking at the graph, it is clear that throughout11PM, the energy use is either at 40 Watt Hour of energy, or near 0. This shows that there is either a high usage, or none at all. Something that could explain this is that either the household would put the dishwasher to use before going to bed, or they wouldn't, and that when the dishwasher is working it's using that precise amount of energy where the dark y line is. The other points could represent when the household is doing additional things in this area of the house, that are taking less / more energy than the constant of a dishwasher. This demonstrates this family generally has a similar schedule for power in this area of the house, with an assumption of putting the dishwasher to use a lot of nights, and the stray points not located near the bulk of the data represents nights their schedule is different. and that various things from this area of the house, such as a microwave or an oven, are being used


###Sasha

```{r}
ggplot(powerC) + 
  geom_jitter(aes(x=Global_active_power, y = Sub_metering_1), color = "royalblue1")+
  geom_jitter(  aes(x=Global_active_power, y = Sub_metering_2), color = "lightpink")+
  geom_jitter( aes(x=Global_active_power, y = Sub_metering_3), color = "slateblue1")+ ylab("Sub Metering") + xlab("Global Active Power")
```


  I graphed the three different Sub Metering as it relates to the Global Active Power. Global Active Power is the household global minute-averaged active power and Sub Metering 1 is the energy used in areas corresponding to the kitchen, usually in the form of a dishwasher, oven or microwave. Sub Metering 2 is the power used in areas corresponding to laundry rooms generally containing a washer, dryer and a refridgerator. Sub Metering 3 corresponds to areas relating to an electric water heater or air conditioner. As seen in the graph, it is clear that submetering 1 and 2 use relatively the same amount of energy while submetering 3 uses way less amount of energy.


###Ping   
I want to explore how the year, month are affecting the meters assumptions. By the first graph I did, I learned that Meter3(water-heater and air-conditioner) is incresing more thtoughout the years, so I did one more step to see how the meter3 usage is changing in 2008.
```{r}
powerC <- powerC %>% mutate(Year = year(dmy(Date)), Month = month(dmy(Date)), Day = day(dmy(Date)))

mySecondData <- powerC %>% filter(powerC$Year != 2006) %>% select(Sub_metering_1,Sub_metering_2,Sub_metering_3,Year, Month)
Jan <- mySecondData %>% filter(mySecondData$Month == 1, mySecondData$Year == 2008)
Feb <- mySecondData %>% filter(mySecondData$Month == 2, mySecondData$Year == 2008)
Mar <- mySecondData %>% filter(mySecondData$Month == 3, mySecondData$Year == 2008)
Apr <- mySecondData %>% filter(mySecondData$Month == 4, mySecondData$Year == 2008)
May <- mySecondData %>% filter(mySecondData$Month == 5, mySecondData$Year == 2008)
Jun <- mySecondData %>% filter(mySecondData$Month == 6, mySecondData$Year == 2008)
Jul <- mySecondData %>% filter(mySecondData$Month == 7, mySecondData$Year == 2008)
Aug <- mySecondData %>% filter(mySecondData$Month == 8, mySecondData$Year == 2008)
Sep <- mySecondData %>% filter(mySecondData$Month == 9, mySecondData$Year == 2008)
Oct <- mySecondData %>% filter(mySecondData$Month == 10, mySecondData$Year == 2008)
Nov <- mySecondData %>% filter(mySecondData$Month == 11, mySecondData$Year == 2008)
Dec <- mySecondData %>% filter(mySecondData$Month == 12, mySecondData$Year == 2008)

janSum <- sum(Jan$Sub_metering_3,na.rm = TRUE)
febSum <- sum(Feb$Sub_metering_3,na.rm = TRUE)
marSum <- sum(Mar$Sub_metering_3,na.rm = TRUE)
aprSum <- sum(Apr$Sub_metering_3,na.rm = TRUE)
maySum <- sum(May$Sub_metering_3,na.rm = TRUE)
junSum <- sum(Jun$Sub_metering_3,na.rm = TRUE)
julSum <- sum(Jul$Sub_metering_3,na.rm = TRUE)
augSum <- sum(Aug$Sub_metering_3,na.rm = TRUE)
sepSum <- sum(Sep$Sub_metering_3,na.rm = TRUE)
octSum <- sum(Oct$Sub_metering_3,na.rm = TRUE)
nowSum <- sum(Nov$Sub_metering_3,na.rm = TRUE)
novSum <- sum(Nov$Sub_metering_3,na.rm = TRUE)
decSum <- sum(Dec$Sub_metering_3,na.rm = TRUE)

Graph <- tibble(Year = c(rep(2008,12)), Total = c(janSum,febSum,marSum,aprSum,maySum,junSum,julSum,augSum,sepSum,octSum,novSum,decSum), Month = c(1:12), Type = rep(3,12))

ggplot(data = Graph) + geom_line(mapping = aes(x = Graph$Month, y = Graph$Total, color = "red")) + labs(x = "Month", y = "Meter3 total", title = "Water-heater & Air-conditioner usage in 2008") + scale_x_continuous(breaks=c(1,3,5,7,9,11))
```

conclusion: I started by analyzing the overall relatiosnhip between years and the trending of Meter1, 2 and 3. I realized Meter3 grows consistently, so I took 2008 to take a deeper look. We can see from the graph that water-heater and air-conditioner usage increases from Feb to Apr. The usage had a huge decreases from June to Aug. This can explain with their summer temperature is comfortable or people travel abroad a lot durin the summers.  