---
title: "Lab8"
author: "R-Kelly"
date: "March 6, 2019"
output: html_document
---

###Team Statement

Overall question we're answering - What's the distribution of Lyme disease in the US?
A backed-up answer to the stated question can help doctors that specialize in the treatment of Lyme disease know where to operate. Knowing the distribution of Lyme 
disease can also potentially inform doctors what the causes of Lyme disease are.

The dataset we used includes the number of cases of Lyme disease broken down by year, and by county of each US state. 
Our domain expert will be medical companies around the US interested in diseases like Lyme Disease.

Conclusion:

Recommendation:


The pretend domain & the reference/link to the data 

conclusion & recommendation


<<<<<<< Updated upstream
```{r, include=FALSE}
library(tidyverse)
lab8data <- read.csv("lab8-file.csv", header = TRUE)
```


=======
###Lauren
```{r}
library(tidyverse)
library(readr)
lime <- read_csv("lab8data.csv")
lime <- lime %>% 
  gather(`Cases2000`, `Cases2001`,`Cases2002`,`Cases2003`,`Cases2004`,`Cases2005`,`Cases2006`, `Cases2007`,`Cases2008`,`Cases2009`,`Cases2010`,`Cases2011`,`Cases2012`,`Cases2013`,`Cases2014`,`Cases2015`,`Cases2016`,`Cases2017`,key = "year", value = "cases")
lime$year[lime$year=='Cases2000'] <- 2000
lime$year[lime$year=='Cases2001'] <- 2001
lime$year[lime$year=='Cases2002'] <- 2002
lime$year[lime$year=='Cases2003'] <- 2003
lime$year[lime$year=='Cases2004'] <- 2004
lime$year[lime$year=='Cases2005'] <- 2005
lime$year[lime$year=='Cases2006'] <- 2006
lime$year[lime$year=='Cases2007'] <- 2007
lime$year[lime$year=='Cases2008'] <- 2008
lime$year[lime$year=='Cases2009'] <- 2009
lime$year[lime$year=='Cases2010'] <- 2010
lime$year[lime$year=='Cases2011'] <- 2011
lime$year[lime$year=='Cases2012'] <- 2012
lime$year[lime$year=='Cases2013'] <- 2013
lime$year[lime$year=='Cases2014'] <- 2014
lime$year[lime$year=='Cases2015'] <- 2015
lime$year[lime$year=='Cases2016'] <- 2016
lime$year[lime$year=='Cases2017'] <- 2017
lime$year <- parse_integer(lime$year)

limetotal<-aggregate(lime['cases'], by= lime['Ctyname'], sum)
limestate<-aggregate(lime['cases'], by= lime['Stname'], sum)

#library(readr)
#uscities <- read_csv("uscitiesv1.4.csv")
#lime$data <- parse_character(lime$Ctyname)

#fips <- read_csv("fipss.csv")


#`+` <- function(e1, e2) {
  ## unary
  #if (missing(e2)) return(e1)
 # if (!is.na(suppressWarnings(as.numeric(e1))) & !is.na(suppressWarnings(as.numeric(e2)))) {
    ## both arguments numeric-like but characters
   # return(base::`+`(as.numeric(e1), as.numeric(e2)))
 # } else if ((is.character(e1) & is.na(suppressWarnings(as.numeric(e1)))) | 
           #  (is.character(e2) & is.na(suppressWarnings(as.numeric(e2))))) {
    ## at least one true character 
 #   return(paste0(e1, e2))
#  } else {
#    ## both numeric
 #   return(base::`+`(e1, e2))
 # }
#}

#uscities$county_name <- uscities$county_name + " County"
#fips$County <- fips$County + " County"

#lime2 <- merge(limetotal, uscities, by.x = "Ctyname", by.y = "county_name", all = TRUE)

#limelocation<- lime2 %>% select(Ctyname, state_name, cases, lat, lng)

#limelocation<- merge(limelocation, fips, by.x="Ctyname", by.y = "County", all = TRUE)


#unique(limelocation, by = "Ctyname")

#limetotal <- limeloacation %>% group_by(Ctyname) %>% summarise(cases = sum(cases)))

#limetotal=tapply(limelocation$cases, cumsum(limelocation$Ctyname==1), FUN=sum, na.rm=TRUE)

#limelocation<-limelocation[!duplicated(limelocation$Ctyname), ]

#limelocation<-na.omit(limelocation)


library(ggplot2)

ggplot(limestate) + geom_bar(mapping=aes(x=Stname))




```

Originally I wanted to plot the cases of lime disease on a map of the United States, and so I merged our original data with data about FIPS for each county, and the longitude and latitude of each county. The plot_usmap function was not working, so I ditched the plot of the US and instead plot the amount of total cases per state. The most cases occured in Pennslyvania and the least in Hawaii.


###Peter
```{r}
datam <- limelocation %>% select(cases, lat)

ggplot(datam) +
  geom_point(aes(x = cases, y = lat))
  
```

###Sasha
```{r setup, include=FALSE}
library(tidyverse)
library(readr)
lymeDisease <- read_csv("lab8data.csv")

lymeDisease <- lymeDisease %>% 
  gather(`Cases2000`, `Cases2001`,`Cases2002`,`Cases2003`,`Cases2004`,`Cases2005`,`Cases2006`, `Cases2007`,`Cases2008`,`Cases2009`,`Cases2010`,`Cases2011`,`Cases2012`,`Cases2013`,`Cases2014`,`Cases2015`,`Cases2016`,`Cases2017`,key = "Year", value = "Cases")
lymeDisease$Year[lymeDisease$Year=='Cases2000'] <- 2000
lymeDisease$Year[lymeDisease$Year=='Cases2001'] <- 2001
lymeDisease$Year[lymeDisease$Year=='Cases2002'] <- 2002
lymeDisease$Year[lymeDisease$Year=='Cases2003'] <- 2003
lymeDisease$Year[lymeDisease$Year=='Cases2004'] <- 2004
lymeDisease$Year[lymeDisease$Year=='Cases2005'] <- 2005
lymeDisease$Year[lymeDisease$Year=='Cases2006'] <- 2006
lymeDisease$Year[lymeDisease$Year=='Cases2007'] <- 2007
lymeDisease$Year[lymeDisease$Year=='Cases2008'] <- 2008
lymeDisease$Year[lymeDisease$Year=='Cases2009'] <- 2009
lymeDisease$Year[lymeDisease$Year=='Cases2010'] <- 2010
lymeDisease$Year[lymeDisease$Year=='Cases2011'] <- 2011
lymeDisease$Year[lymeDisease$Year=='Cases2012'] <- 2012
lymeDisease$Year[lymeDisease$Year=='Cases2013'] <- 2013
lymeDisease$Year[lymeDisease$Year=='Cases2014'] <- 2014
lymeDisease$Year[lymeDisease$Year=='Cases2015'] <- 2015
lymeDisease$Year[lymeDisease$Year=='Cases2016'] <- 2016
lymeDisease$Year[lymeDisease$Year=='Cases2017'] <- 2017
lymeDisease$Year <- parse_integer(lymeDisease$Year)

lymeDisease$STCODE <- NULL
lymeDisease$CTYCODE <- NULL


ggplot(data = lymeDisease, aes(x=Year, y=Cases, color = Stname)) +
  geom_bar(stat="identity")
```

For my graph, I wanted to show the prevalence of lyme diseases per state per year. To draw a conclusion from our original dataset, it was necessary to gather the case columns and make two new coloumns for Case and Year. I then chose to get rid of columns that were redundant namely the state code and county code column. I made the graph a bar plot with the X axis being characterized by year while the Y axis was characterized by the amount of cases. From the graph, the amount of cases of lyme disease is increasing yearly but there is not a clear prevalence of lyme disease in a particular state.



### Gregor

All of Tidying / Arranging Data Set

```{r}
lime <- limeDisease %>% 
  gather(`Cases2000`, `Cases2001`,`Cases2002`,`Cases2003`,`Cases2004`,`Cases2005`,`Cases2006`, `Cases2007`,`Cases2008`,`Cases2009`,`Cases2010`,`Cases2011`,`Cases2012`,`Cases2013`,`Cases2014`,`Cases2015`,`Cases2016`,`Cases2017`,key = "Year", value = "Cases")
lime$Year[lime$Year=='Cases2000'] <- 2000
lime$Year[lime$Year=='Cases2001'] <- 2001
lime$Year[lime$Year=='Cases2002'] <- 2002
lime$Year[lime$Year=='Cases2003'] <- 2003
lime$Year[lime$Year=='Cases2004'] <- 2004
lime$Year[lime$Year=='Cases2005'] <- 2005
lime$Year[lime$Year=='Cases2006'] <- 2006
lime$Year[lime$Year=='Cases2007'] <- 2007
lime$Year[lime$Year=='Cases2008'] <- 2008
lime$Year[lime$Year=='Cases2009'] <- 2009
lime$Year[lime$Year=='Cases2010'] <- 2010
lime$Year[lime$Year=='Cases2011'] <- 2011
lime$Year[lime$Year=='Cases2012'] <- 2012
lime$Year[lime$Year=='Cases2013'] <- 2013
lime$Year[lime$Year=='Cases2014'] <- 2014
lime$Year[lime$Year=='Cases2015'] <- 2015
lime$Year[lime$Year=='Cases2016'] <- 2016
lime$Year[lime$Year=='Cases2017'] <- 2017
lime$Year <- parse_integer(lime$Year)

library(plyr)
colnames(lime)[2] <- "State"
colnames(lime)[3] <- "State Code"
lime$Ctyname <- NULL
lime$CTYCODE <- NULL
lime2 <- ddply(lime,"State",numcolwise(sum))
lime2 <- mutate(lime2, HighState = ifelse(State == "Colorado" | State == "Wyoming" | State == "Utah" | State == "New Mexico" | State == "Nevada" | State == "Idaho" | State == "Arizona" | State == "Oregon" | State == "Hawaii", "Yes", "No"))

lime3 <- filter(lime2, HighState == "No")
lime4 <- filter(lime2, HighState == "Yes")

```



```{r}
ggplot(lime3, aes(x = State, y = Cases)) + geom_histogram(stat = "identity") + coord_flip() + ylab("Cases of Lyme Disease from 2000-2017 in states with low average elevation") + ylim(0, 100000)

ggplot(lime4, aes(x = State, y = Cases)) + geom_histogram(stat = "identity") + coord_flip() + ylab("Cases of Lyme Disease from 2000-2017 in states with high average elevation") + ylim(0, 100000)
```


I wanted to determine whether there was a correlation between number of total cases of Lyme disease throughout these 17 years and certain areas of the US. Initially, I tidied the raw set so that years would be under one column. I then had to sum up the cases for all of the years, and to do this I used a function from the "plyr" package, called ddply. In order to be able to see the difference in cases between different areas in the US, I created another variable called "High State" in my data frame using the mutate function. Within the mutate I used the base R function ifElse, so that I could make this variable on the condition of the state having a higher elevation.

My data/graphs show that there are much less cases in the states that have a higher average elevation. My first graph shows the number of cases of Lyme disease in states that have a lower average elevation, and it is clear that there are many states that have a very high number, reaching up to almost 100,000. When looking at the second graph of the number of cases in states with a higher average elevation, it's apparent that the average is much lower; none of the states reach close to the 1000s. I used the same y value limit so that the difference could be observed. This shows a clear relationship between elevation and number of Lyme cases.

To do this analysis, I did need to use outside information to find a ranking of which states had the highest elevation. If the original data set included a variable for elevation, a geom_smooth function could be used with elevation vs. number of cases, and it would show a clear linear relationship. 

### Ping

library(tidyverse)
library(readr)
library(ggplot2)
Lyme_00_17 <- read_csv("lab8data.csv")
cases_by_state = lapply(split(Lyme_00_17[,c('Cases2000', 'Cases2001','Cases2002','Cases2003','Cases2004','Cases2005','Cases2006','Cases2007','Cases2008','Cases2009','Cases2010','Cases2011','Cases2012','Cases2013','Cases2014','Cases2015','Cases2016','Cases2017')], Lyme_00_17$Stname), colSums)
cases_by_state = do.call(rbind, cases_by_state)
cases_by_state = data.frame(cases_by_state)
cases_by_state$State <- as.character(rownames(cases_by_state))
cases_by_state <- dplyr::rename(cases_by_state, '2000' = "Cases2000",'2001' = "Cases2001", '2002' = "Cases2002", '2003' = "Cases2003", '2004' = "Cases2004", '2005' = "Cases2005", '2006' = "Cases2006", '2007' = "Cases2007", '2008' = "Cases2008", '2009' = "Cases2009", '2010' = "Cases2010", '2011' = "Cases2011", '2012' = "Cases2012", '2013' = "Cases2013", '2014' = "Cases2014", '2015' = "Cases2015", '2016' = "Cases2016", '2017' = "Cases2017" )
firstdata <- cases_by_state  %>% select('2000','2003','2006','2009','2012', '2015',State)
graphdata <- firstdata %>% filter ( State == "Maine"| State == "Colorado"|State == "California" | State == "Connecticut" | State == "New Jersey" | State == "Massachusetts" | State == "Minnesota" | State == "Wyoming" | State == "New York")
graphdata_melt = reshape2::melt(graphdata)
ggplot(data = graphdata_melt, aes(x = variable, y = value, color = State)) + geom_line()+ facet_wrap(~State) + theme(axis.text.x = element_text(angle = 90))




>>>>>>> Stashed changes