---
title: "Lab8"
author: "R-Kelly"
date: "March 6, 2019"
output: html_document
---

###Team Statement

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<<<<<<< Updated upstream
```{r, include=FALSE}
library(tidyverse)
Lyme_00_17 <- read.csv("lab8-file.csv", header = TRUE)
```


=======
###Lauren
```{r}
library(tidyverse)
library(readr)
lime <- read_csv("lab8data.csv")
View(lab8data)
lime <- lime %>% 
  gather(`Cases2000`, `Cases2001`,`Cases2002`,`Cases2003`,`Cases2004`,`Cases2005`,`Cases2006`, `Cases2007`,`Cases2008`,`Cases2009`,`Cases2010`,`Cases2011`,`Cases2012`,`Cases2013`,`Cases2014`,`Cases2015`,`Cases2016`,`Cases2017`,key = "year", value = "cases")
lime$year[lime$year=='Cases2000'] <- 2000
lime$year[lime$year=='Cases2001'] <- 2001
lime$year[lime$year=='Cases2002'] <- 2002
lime$year[lime$year=='Cases2003'] <- 2003
lime$year[lime$year=='Cases2004'] <- 2004
lime$year[lime$year=='Cases2005'] <- 2005
lime$year[lime$year=='Cases2006'] <- 2006
lime$year[lime$year=='Cases2007'] <- 2007
lime$year[lime$year=='Cases2008'] <- 2008
lime$year[lime$year=='Cases2009'] <- 2009
lime$year[lime$year=='Cases2010'] <- 2010
lime$year[lime$year=='Cases2011'] <- 2011
lime$year[lime$year=='Cases2012'] <- 2012
lime$year[lime$year=='Cases2013'] <- 2013
lime$year[lime$year=='Cases2014'] <- 2014
lime$year[lime$year=='Cases2015'] <- 2015
lime$year[lime$year=='Cases2016'] <- 2016
lime$year[lime$year=='Cases2017'] <- 2017
lime$year <- parse_integer(lime$year)

limetotal<-aggregate(lime['cases'], by= lime['Ctyname'], sum)

library(readr)
uscities <- read_csv("uscitiesv1.4.csv")
lime$data <- parse_character(lime$Ctyname)

`+` <- function(e1, e2) {
  ## unary
  if (missing(e2)) return(e1)
  if (!is.na(suppressWarnings(as.numeric(e1))) & !is.na(suppressWarnings(as.numeric(e2)))) {
    ## both arguments numeric-like but characters
    return(base::`+`(as.numeric(e1), as.numeric(e2)))
  } else if ((is.character(e1) & is.na(suppressWarnings(as.numeric(e1)))) | 
             (is.character(e2) & is.na(suppressWarnings(as.numeric(e2))))) {
    ## at least one true character 
    return(paste0(e1, e2))
  } else {
    ## both numeric
    return(base::`+`(e1, e2))
  }
}

uscities$county_name <- uscities$county_name + " County"

lime2 <- merge(limetotal, uscities, by.x = "Ctyname", by.y = "county_name", all = TRUE)

limelocation<- lime2 %>% select(Ctyname, cases, lat, lng)

#unique(limelocation, by = "Ctyname")

#limetotal <- limeloacation %>% group_by(Ctyname) %>% summarise(cases = sum(cases)))

#limetotal=tapply(limelocation$cases, cumsum(limelocation$Ctyname==1), FUN=sum, na.rm=TRUE)

limelocation<-limelocation[!duplicated(limelocation$Ctyname), ]
limelocation<-na.omit(limelocation)



```


```{r}

```


# grab a center/zoom map and compute its bounding box
gc <- geocode("white house, washington dc")
map <- get_map(gc)
(bb <- attr(map, "bb"))
(bbox <- bb2bbox(bb))
# use the bounding box to get a stamen map
stamMap <- get_stamenmap(bbox)
ggmap(map) +
geom_point(
aes(x = lon, y = lat),
data = gc, colour = "red", size = 3
)
ggmap(stamMap) +
geom_point(
aes(x = lon, y = lat),
data = gc, colour = "red", size = 3
)
###Peter
```{r}
datam <- limelocation
datam$lat <- as.integer(datam$lat)
datam$cases <- as.numeric(datam$cases)
c <- ggplot(datam, aes(x = cases, y = lat)) +
  geom_smooth( color = 'red')
```
>>>>>>> Stashed changes