---
title: "Lab15"
author: "R-Kelly"
date: "May 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#libraries
library(tidyverse)
library(dplyr)
library(modelr)
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
#File upload:
#install.packages("ggpubr")
library(readr)
library(tidyverse)
library(modelr)
#library(ggpubr)
team_info<- read_csv("team_info.csv")
player_info<- read_csv("player_info.csv")
game_teams_stats<-read_csv("game_teams_stats.csv")
game_skater_stats<-read_csv("game_skater_stats.csv")
game_goalie_stats<-read_csv("game_goalie_stats.csv")
game<-read_csv("game.csv")
```


> Team question: What makes a hockey team successful?

###Individual section
##Ping
> I choose to examine how goalies' performanes affect the winning ratio of the team. A goalie is one of the most important people in the team, so their performances definitely decided the scores of the team. By exploring how it is affecting the team's scores, we can see does golaies' performances matter to a hockey  team's success or not.
> I will use two different methods to examine the result. The first is using ggplot to plot a point graph to see how the trending is going in winning teams and lossing teams. And then I will use permutation correlation test to see what is more significant.

```{r}
game_team <- game_teams_stats
goalie <- game_goalie_stats

combined <- full_join(game_team, goalie, by = c("game_id", "team_id"))
combined <- na.omit(combined)

combined <- combined %>% group_by(team_id) %>% select(game_id, team_id, won, player_id, saves, powerPlaySaves, shortHandedSaves, evenSaves, savePercentage, powerPlaySavePercentage, evenStrengthSavePercentage)
combined$team_id <- as.factor(combined$team_id)

ggplot(data = combined, mapping = aes(x = saves, y = powerPlaySaves, color = team_id)) + geom_point() + facet_wrap(~won)
ggplot(data = combined, mapping = aes(x = saves, y = evenSaves, color = team_id)) + geom_point() + facet_wrap(~won)
```

> We can see from the graph abover, the teams won have a more densed graph at the middle. So we can say, for the teams that won, they have more saves in gernal, and have a higher amounts of power play saves also. We can see the same thing for even saves as well. They have a more densed distribution and more total number of saves and even saves.

```{r}
perm_cor <- function(perms = 100, year, mean)
{
 vector <- c(rep(0,perms))

 # Loop throught number of permutations
 for (i in c(1:perms))
 {
   # Step 2:
   # Randomly mix up the values in the vector "y"
   mean <- sample(mean, length(mean))

   # Step 3:
   # Compute the correlation between x and the randomly mixed
   # up y-vector. Store this value in the vector from step 1.
   vector[i] <- cor(mean, year)
 }

 # Step 4:
 # Return new updated vector, created in step 1
 return(vector)
}
```
```{r}
winning <- combined %>% filter(won == "TRUE")
loss <- combined %>% filter(won == "FALSE")
cor_save <- perm_cor(perms = 1000, combined$powerPlaySaves, combined$saves)
cor_save <- data.frame(cor_save)
true_corr_save1 <- cor(winning$powerPlaySaves, winning$saves)
true_corr_save2 <- cor(loss$powerPlaySaves, loss$saves)

ggplot(cor_save) + geom_density(aes(x = cor_save))  + geom_vline(aes(xintercept=true_corr_save1), linetype="dashed", size=1, color = "blue") +    
geom_vline(aes(xintercept=true_corr_save2), linetype="dashed", size=1, color = "red") +
  ggtitle("The correlation between power play saves and total saves")

cor_even <- perm_cor(perms = 1000, combined$evenSaves, combined$saves)
cor_even <- data.frame(cor_even)
true_corr_save3 <- cor(winning$evenSaves, winning$saves)
true_corr_save4 <- cor(loss$evenSaves, loss$saves)

ggplot(cor_even) + geom_density(aes(x = cor_even)) +    geom_vline(aes(xintercept=true_corr_save3), linetype="dashed", size=1, color = "blue") +    
geom_vline(aes(xintercept=true_corr_save4), linetype="dashed", size=1, color = "red") +
  ggtitle("The correlation between even play saves and total saves") 

```

> The red line represents the correlation for lossing teams, and blue line represents the correlation for winning teams. And we can see that lossing teams have a more significnat result.
> As a result, we can say that a goalie's performance affect a team's winning ratio. So one of the factors that makes a team more successful is the goalie's performance.







##Gregor
```{r}
#cleaning data sets
game$venue_link <- NULL
game$venue_time_zone_id <- NULL
game$venue_time_zone_offset <- NULL
game$venue_time_zone_tz <- NULL
game$date_time < NULL
#game has date, type of game, each team id, home and away goals, result and whether it came in overtime or regulation, home rink side and venue
```

```{r}
FinalDataGregor <- game_teams_stats
FinalDataGregor <- game_teams_stats %>% left_join(team_info, by = "team_id")
FinalDataGregor$franchiseId <- NULL
FinalDataGregor$link <- NULL
FinalDataGregor$team_id <- NULL
#joining so that I have proper team name, and additional clean up so that looking at the dataset in the table is a little more organized
```

```{r}
#adding a few useful columns
FinalDataGregor <- FinalDataGregor %>% mutate(ShotPCT = (goals/shots)*100 , TakeawaysGiveawaysDiff = takeaways - giveaways)

#creating new data set with huge summarise, obtaining season averages for various statistics for each team
SeasonAverages <- FinalDataGregor %>% select(teamName, won, goals, shots,ShotPCT, pim, faceOffWinPercentage,hits,powerPlayOpportunities,powerPlayGoals, hits) %>% group_by(teamName) %>% summarise_all(mean)
```

```{r, include = FALSE}

GatherCorrelationalValues <- function(df){
  
  array1 <- c()
  
  for (i in 3:ncol(df)){
    array1[i] <- cor(SeasonAverages[[i]], SeasonAverages$won)
    
  }
  
  array2 <- c()
  
  for (i in 3:ncol(df)){
    array2[i] <- colnames(SeasonAverages)[i]
  }
  
  df <- tibble(ColName = c(array2), Correlation = c(array1))
  
  return(df)
  
}
#obtain dataset so that I can graph correlational values of different columns with WinPCT
CorrelationalData <- GatherCorrelationalValues(SeasonAverages)
```

Through all of this data modifying/collection, I am able to graph my first practical findings.

The following graph demonstrates the various statistics, and the correlation they have with win percentage (these are over the course of the season, so there is tons of data behind these numbers, this was done by doing the summarise dplyr function a little farther above)

```{r}
CorrelationalData <- na.omit(CorrelationalData)
ggplot(data = CorrelationalData, aes(x = ColName,y =Correlation)) + geom_bar(stat = "identity", fill = "steel blue") + coord_flip() + ggtitle("Correlation between various NHL statistics with Win Percentage")
```

By looking at the graph above, there are some things that are obvious, but also a couple that stand out as well. Clearly, win percentage will have a perfect correlation with win percentage, and number of goals also will have a very high correlation with win percentage, since in order to win you must have more goals than the other team. Something maybe not quite as obvious is that Shot Percentage, the percentage of shots that result to goals, has a higher correlation to wins than shots does, so perhaps teams that focus on getting better shots rather than shooting lots of shots are more successful. Penalty in minutes is a negative correlation because generally the more time a team has players in the penalty box the greater the disadvantage they create for themselves. However, since it's such a low correlation, one could say that maybe penalties aren't such a big hinder to the game. 

One of the higher correlations is Power Play Goals. Let's create a model between power play goals and win percentage.



```{r}
GregorModel <- lm(powerPlayGoals~won, SeasonAverages)

SeasonAverages <- add_residuals(SeasonAverages, GregorModel)

ggplot(SeasonAverages, aes(teamName, resid)) + geom_point(color = "red") + coord_flip()
```

By looking at the graph above, the range of the residuals is very low. That means there are few teams that stray far from the model of power play goals leading to a greater win percentage. 

Therefore, it is relatively safe to say that a team that is effecient at scoring power play goals tends to be a successful team. 


Checking to see how this compares with the top teams in the NHL from this dataset...


```{r}
PowerPlayGoalsLeagueAverage = mean(SeasonAverages$powerPlayGoals)
SeasonAverages <- SeasonAverages %>% mutate(PowerPlayGoalDiffFromAvg = powerPlayGoals - PowerPlayGoalsLeagueAverage)
ggplot(SeasonAverages, aes(x = teamName, y = PowerPlayGoalDiffFromAvg)) + geom_bar(stat = "identity") + coord_flip() + ggtitle("Residual of model: Power Play Goals and Win PCT")

```


The teams that have the greatest positive difference from the league average of Power Play Goals are the Capitals and Penguins, 2 of the top 3 teams in the league according to Win Percentage, and the team that has the greatest negative difference are the Panthers, who are bottom 5 in win percentage. 

##Sasha

```{r}
stats <- game_teams_stats
options(na.action = na.warn)
away <- lm(takeaways ~ giveaways, data = stats)
coef(stats)

stats <- stats %>% 
  add_residuals(away)

away1 <- filter(stats, resid >= 0)

ggplot(away1) +
  geom_histogram(aes(x = takeaways, fill = settled_in))+
  geom_histogram(aes(x = resid, fill = "Residual", alpha = 0.8 ))+
  facet_grid(~won)+
  labs(
    title = "Takeaways",
    x = "Goals",
    y = "Takeaways"
  )

ggplot(away1) +
  geom_histogram(aes(x = giveaways, fill = settled_in))+
  geom_histogram(aes(x = resid, fill = "Residual", alpha = 0.8 ))+
  facet_grid(~won)+
  labs(
    title = "Giveaways",
    x = "Goals",
    y = "Giveaways"
  )

```
My individual subquestion dealt with the relationship between takeaways and giveaways
 and how they relate to wins and losses. This is important because it will reinforce 
 the idea that puck possesion is key to winning and maybe even suggest that the "Dump 
 and Chase" style of playing hockey is not strategically successful.

I worked with models and transformations to data. 

Winning teams have more takeaways per game than losing teams, this can be seen in both 
the regular data and the found residual model. Games that end in a shoot out are not 
especially dependant on this finding however, but, games that are setteled in regulation 
and over time rely heavily on this finding.On the flip side, in terms of giveaways, the 
winning teams tend to have more giveaways. This is interesting because in ice hockey, 
puck possession is key to scoring goals, and one would imagine that less giveways would 
be a surefire sign of success. So inconclusion, the number of giveaways a team has is 
largely irrelevant, and it is way more important to increase the number of takeaways. 
The relationship between giveaways and game length are similar to that of takeaways, 
where it is pretty insignificant in a shoot out but relatively significant in over time and 
regulation.

Our overall question was asking what makes a hockey team successful, from the data we can 
see that takeaways are the name of the game. No matter how many times you lose puck 
possession, if you can increase takeaways, there is a larger chance those will result in 
goals and thus a win!

I don't really think I had a six month goal to begin with, however, my current six month 
goal is to spend a lot of time working in R to solidify my skills and knowledge. My 5 year 
goal is to graduate and find a job probably in the financial sector. I learned R, for one, 
and I learned a lot about data science in general. It was really rewarding to get have break 
throughs with tough code. I would tell myself to start to read the textbook more closely,
 keep working at it even when you're discouraged and stop being unorganized! 


##Peter
```{r, message = FALSE, warning = FALSE}
data <- game_teams_stats
dtest <- function(inDat, z){
  vec <- numeric(2)
  for (i in 1:(length(inDat[[1]])-1)){if(i%%2 != 0)next else if(inDat[[z]][[i]]>inDat[[z]][[i+1]] && inDat[[4]][[i]] == TRUE){vec[[2]] <- vec[[2]] + 1}else{vec[[1]] <- vec[[1]] + 1}}
  return(vec)
}
names <- colnames(data)
bigRes <- numeric(7)
for (i in 8:15){
  temp <-dtest(data, i)
  d <- (temp[[1]])/(temp[[1]]+temp[[2]])
  print(names[[i]])
  print(d)

  bigRes[[i-7]] <- d
}

plot(bigRes)
```
For my individual section, I thought I would examine the relationship between individual game measures and winning teams. I wanted to know which individual factor indicates a winning team most often. 

I did this by finding how often each stat winner (team with more recorded events in the categroy) won the game. I thought this would 
show what measures best predict a winning team, and the relative predictive ability of each.

To do this analysis, I used a number of For loops, conditionals, and a function. One for loop is non-linear, as observations came in pairs in the dataset. 
This was a challenging task, as I had previously not modified for loop iteration in R, but I was happy to figure it out. 

My results show a higher than average predictive ability for Power Play shots and Goals. This makes sense, as these stats describe decisive scoring moments, but it shows their relative importance over other factors. No factor scored higher than .79, which shows that no factor is a perfect predictor of team success. 

After this course, my six month goal remains the same, but with much more exprience to draw from. I wanted to learn more skills to help apply my ideas and I feel much more equipped now than at the beginning of the semester. I hope to keep attempting challenging problems, and improving my analytical abilites. 
My five year goal remains the same, but now with more context as to what being a data scientist is. If I could give myself advice at the beginning of the semester, I would encourage myself to attempt the hardest analysis I could each lab, as it would have helped my skillset even more. 


##Lauren